Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-09-06 paper-correlations.tex
Remove self-repeating passages.

\begin{abstract}
Correlation functions are becoming one of the major tools for quantification of
structural information that is usually represented as material's 2D or 3D image.
In this paper we introduce \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on imaging input data. Images include both binary and multi-phase
representations.  Our code is capable of evaluating two-point probability $S_2$,
cluster $C_2$, lineal-path $L_2$, surface-surface $F_{ss}$, surface-void
$F_{sv}$, pore-size $P$ and chord-length $p$ functions on both CPU and GPU
architectures.  There possible, we presented two types of computations: full
correlation map (correlations of each point with other pints on the image, that
also allows obtaining ensemble averaged CF) and directional correlation
functions (currently in major orthogonal and diagonal directions).  Such an
implementation allowed for the first time to assemble a completely free solution
to evaluate correlation functions under any operating system with well
documented application programming interface (API). Our package includes
automatic tests against analytical solutions that are described in the paper. We
measured wall times for all CPU and GPU implementations and as a rule of thumb
full correlation maps on GPU are faster than directional scans using CPU due to
the usage of fast Fourier transform. However, full maps require more RAM and,
thus, are limited to available GPU RAM resources. On the other hand, directional
CFs are memory efficient and can be evaluated for huge datasets - this way they
are the first candidates for structural data compression of feature
extraction. The package itself is available through Julia package ecosystem and
on GitHub (https://github.com/fatimp/CorrelationFunctions.jl), the latter source
also contains documentation and additional helpful resources such as tutorials.
We believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications. We also present some example as
applied to ceramic, soil composite and oil-bearing rock samples based on their
3D X-ray tomography and 2D scanning electron microscope images. Finally, we
conclude our paper with discussion of possible ways to further improve presented
computational framework.
\end{abstract}

@@

\section{Summary}
\label{sec:summary}
In this paper we introduced \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on 2D and 3D input images. Our code is capable of evaluating two-point
probability, cluster, lineal-path, surface-surface, surface-void, pore-size and
chord-length functions on both CPU and GPU architectures. There possible, we
presented two types of computations: full correlation map (correlations of each
point with other pints on the image, that also allows obtaining ensemble
averaged CF) and directional correlation functions (currently in major
orthogonal and diagonal directions). Such an implementation allowed for the
first time to assemble a completely free solution to evaluate correlation
functions under any operating system with well documented API. Our package
includes automatic tests against analytical solutions that are described in this
paper. We measured wall times for all CPU and GPU implementations and as a rule
of thumb full correlation maps are faster than directional scans because
calculation of correlation maps applies fast Fourier transform to the whole
input array while, when calculating scans, FFT needs to be applied to each scan
separately. Calculation on GPU shows performance increase when compared to CPU,
as expected.  However, full maps require more RAM and, thus, are limited to
available RAM resources. On the other hand, directional CFs are memory efficient
and can be evaluated for huge datasets - this way they are the first candidates
for structural data compression of feature extraction. The package itself is
available through Julia package ecosystem and on GitHub
(https://github.com/fatimp/CorrelationFunctions.jl), the latter source also
contains documentation and additional helpful resources such as tutorials.  We
believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications.

>>>>

\begin{abstract}
Correlation functions are becoming one of the major tools for quantification of
structural information that is usually represented as material's 2D or 3D image.
In this paper we introduce \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on imaging input data. Images include both binary and multi-phase
representations.  Our code is capable of evaluating two-point probability $S_2$,
cluster $C_2$, lineal-path $L_2$, surface-surface $F_{ss}$, surface-void
$F_{sv}$, pore-size $P$ and chord-length $p$ functions on both CPU and GPU
architectures.  There possible, we presented two types of computations: full
correlation map (correlations of each point with other pints on the image, that
also allows obtaining ensemble averaged CF) and directional correlation
functions (currently in major orthogonal and diagonal directions).  Such an
implementation allowed for the first time to assemble a completely free solution
to evaluate correlation functions under any operating system with well
documented application programming interface (API). Our package includes
automatic tests against analytical solutions that are described in the paper. We
measured wall times for all CPU and GPU implementations and as a rule of thumb
full correlation maps on GPU are faster than other methods. However, full maps
require more RAM and, thus, are limited to available RAM resources. On the other
hand, directional CFs are memory efficient and can be evaluated for huge
datasets -- this way they are the first candidates for structural data
compression of feature extraction. The package itself is available through Julia
package ecosystem and on GitHub
(https://github.com/fatimp/CorrelationFunctions.jl), the latter source also
contains documentation and additional helpful resources such as tutorials.  We
believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications. We also present some example as
applied to ceramic, soil composite and oil-bearing rock samples based on their
3D X-ray tomography and 2D scanning electron microscope images. Finally, we
conclude our paper with discussion of possible ways to further improve presented
computational framework.
\end{abstract}

@@

\section{Summary}
\label{sec:summary}
In this paper we introduced \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on 2D and 3D input images. Our code is capable of evaluating two-point
probability, cluster, lineal-path, surface-surface, surface-void, pore-size and
chord-length functions on both CPU and GPU architectures. There possible, we
presented two types of computations: full correlation map (correlations of each
point with other pints on the image, that also allows obtaining ensemble
averaged CF) and directional correlation functions (currently in major
orthogonal and diagonal directions). Such an implementation allowed for the
first time to assemble a completely free solution to evaluate correlation
functions under any operating system with well documented API. Our package
includes automatic tests against analytical solutions that are described in this
paper. We measured wall times for all CPU and GPU implementations and as a rule
of thumb full correlation maps are faster than directional scans because
calculation of correlation maps applies fast Fourier transform to the whole
input array while, when calculating scans, FFT needs to be applied to each scan
separately. Calculation on GPU shows performance increase when compared to CPU,
as expected.  However, full maps require more RAM and, thus, are limited to
available RAM resources. On the other hand, directional CFs are memory efficient
and can be evaluated for huge datasets -- this way they are the first candidates
for structural data compression of feature extraction. The package itself is
available through Julia package ecosystem and on GitHub
(https://github.com/fatimp/CorrelationFunctions.jl), the latter source also
contains documentation and additional helpful resources such as tutorials.  We
believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications.
